|

### 2020-03-05 analysis/paper/manuscript.Rmd
Draft responses up to minor comments on NCAM explorer

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples). While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019]. We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document. We also used the crosstalk package to link the data displayed across multiple plots. Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data. The intent of this tool was not to replace detailed documentation on the model, however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations.

@@

The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations. Consequently, there exists the possibility of generalizing these tools to simplify their application to other regions. Region specific data management and modelling approaches, however, will surely complicate matters as tools built for one region may be difficult to apply in another region. Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications to be applied to another model being developed. Nevertheless, considerable chunks of code could be recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard. This effort was one step forward in improving the generality of the tool and another step, which is actively being worked on by a colleague in IFREMER, is to modify the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014]. Though more time is required to make the dashboard code more general as the developer must ensure the code is robust to a wider range of scenarios, substantial value could be added by making it more widely applicable. Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g. StoX software; @johnsen2019]. These are, of course, early days in the development and application of interactive tools to stock assessment. Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards. Regardless of the approach, it is important that the path taken is transparent and reproducible.

>>>>

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples). While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore data and models in ways not amenable to static formats and, in the process, these heuristic improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive "dashboard", called NCAM explorer, was developed for the 2018 assessment of Northern cod [@dwyer2019]. We used R-based packages (Table \@ref(tab:pkg)) to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard package to group interactive plotly-based visuals into a dynamic document. We also used the crosstalk package to link the data displayed across multiple plots. Via R Markdown, the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data. The intent of this tool was not to replace detailed documentation on the model [i.e. @cadigan2016; @dwyer2019], however, it was built to serve as an interactive alternative to a standard presentation for communicating the latest results to an audience with a working knowledge of the model.

The NCAM dashboard (Supplement 2) contains a series of pages, the first of which provides terse point-form background on the model (page named "Background"). Subsequent pages provide a series of diagnostic plots for assessing model fits to catch ("Catch"), survey ("RV survey", "SN survey" and "SS survey"; accessed from the "Surveys" drop-down menu), and tagging ("Tagging") data. For instance, the "RV survey" page includes plots of observed and predicted values of mean numbers per tow captured in the research vessel survey (Figure \@ref(fig:rv)). The dashboard also includes pages focused on model estimates such as catchability and selectivity ("Catchability"), stock size and vital rates ("Trends"; Figure \@ref(fig:trends)), and stock productivity ("Productivity"). Finally, some results from a retrospective analyses are included under the "Retro" page, trends from different models are compared under the "Comps" page, details on the projections are accessed from the "Projections" drop-down menu ("Assumptions", "Past projections", "Retro projections", and "Results" pages), and key inputs and outputs are accessed from "Tables" drop-down menu ("Inputs", "Settings", "Outputs" pages). The plots and tables included in the dashboard are similar to those typically presented at assessment meetings and in research documents. However, there are two key benefits of this approach over the modus operandi of producing static documents and slides.  First, interactive plots nested in a dashboard permit relatively easy and efficient access to the details as it replaces scrolling through tens, if not hundreds, of pages or slides with mouse-clicks across pages holding data-rich illustrations (i.e. both broader patterns and finer details in the data are accessible via zooming and tooltips).  Second, the automated nature of the dashboard circumvents the monotonous, time-consuming and error-prone task of copying and pasting figures, tables and values into documents and slides, although, we note that this point is not exclusive to the proposed approach (e.g. slide shows can easily be made using R markdown). Both benefits expedite the process of exploring a range of model configurations as the automated output facilitate quick views of standard diagnostics and the interactive plots facilitate detailed explorations and comparisons of models with different configurations. For instance, explorations of the residual plots in the dashboard allowed the authors to quickly identify conflicts between two surveys used in the model. Potential solutions to this problem were then explored by modifying the configuration of the model and these changes were quickly evaluated using a dashboard that is produced in seconds.

@@

The interactive tools highlighted here were custom built to solve data visualization challenges that are common to many fisheries management organizations. Consequently, there exists the possibility of generalizing these tools to simplify their application to other regions. Region specific data management and modelling approaches, however, will surely complicate matters as tools built for one region may be difficult to apply in another region. Even within our own region, different approaches and needs meant that the code behind NCAM explorer (Supplement 2) required modifications to be applied to another model being developed. Nevertheless, considerable chunks of code could be recycled and, in the process, new features were added to allow the visualization of multiple model configurations within the one dashboard. This effort was one step forward in improving the generality of the tool and another step, which is actively being worked on by a colleague in IFREMER, is to modify the tool to visualize output from SAM, a state-space stock assessment model widely used to assess ICES stocks [@nielsen2014]. Though more time is required to make the dashboard code more general as the developer must ensure the code is robust to a wider range of scenarios, substantial value could be added by making it more widely applicable. Likewise, the concept behind RStrap Explorer and the tagging tool could be applied to other regions and the most efficient approach may be to build upon existing general frameworks [e.g. StoX software; @johnsen2019]. These are, of course, early days in the development and application of interactive tools to stock assessment. Because the tools we highlight are themselves general, flexible and accessible, we suspect that bespoke applications will continue to be developed in parallel with more general stock assessment dashboards. Regardless of the approach, one should strive to make the path taken transparent and reproducible.
